\vspace{-2.0cm}The effect of nitrogen application time on infection and development of the Wheat Streak Mosaic Virus (WSMV)

### 2015-10-26 leslieandkenny.Rnw
Kenny is feeling humbled. Fixed model code and corrected interpretations to match.

<<fakedata,echo=FALSE>>=
# Create some fake data with no differences anywhere
set.seed(983724)
fert.sim1 <- data.frame(expand.grid(
  "status" = c("CNTL", "INOC"),
  "n.application" = c("FALL", "EARLY SPRING", "LATE SPRING"),
  "variety" = c("Var1", "Var2", "Var3", "Var4", "Var5"),
  "block" = LETTERS[1:6],
  "total" = 30),
  "infected" = rnbinom(180, 4, 7/9))
#  "infected" = sample(fert2$infected, 180, replace = TRUE))
@

@@

<<comparison, echo=FALSE, message=FALSE, results='asis'>>=
glmQBin.noSNMS <- glm(cbind(infected, total) ~ block*variety +
                        variety*n.trt*status2,
               family = quasibinomial, data = subset(fert2, variety != "SNMS"))
print(xtable(summary(glmQBin.noSNMS)$coefficients[39:44,], caption="Estimates, standard errors, and tests for three way interaction coefficients in the quasibinomial model.", label="qbin"), caption.placement="top", size="footnotesize", table.placement = getOption("xtable.table.placement", "H"))

noSNMS3way <- glmer(cbind(infected, total) ~ variety*n.trt*status2 +
                    (1|block/row), subset = variety != "SNMS",
                    control = glmerControl(optimizer = "bobyqa"),
                    family = binomial, data = fert2)

@@

<<startingmodel,warning=FALSE,results='hide',cache=TRUE>>=
glm3way <- glm(cbind(infected, total) ~ block*variety + variety*n.trt*status2,
               family = quasibinomial, data = fert.sim1)
@

@@

<<novar1,warning=FALSE,results='hide',cache=TRUE>>=
noVar1 <- glm(cbind(infected, total) ~ block*variety + variety*n.trt*status2,
              family = quasibinomial, data = fert.sim1, subset = variety!="Var1")
@

@@

<<separatemodels,warning=FALSE,results='hide',cache=TRUE>>=
onlyINOC <- glm(cbind(infected, total) ~ block*variety + variety*n.trt,
                family = quasibinomial, data = fert.sim1, subset = status=="INOC")

onlyCNTL <- glm(cbind(infected, total) ~ block*variety + variety*n.trt,
                family = quasibinomial, data = fert.sim1, subset = status=="CNTL")
@

<<models5,echo=FALSE,warning=FALSE,message=FALSE,results='hide',cache=TRUE>>=
glm3way2 <- glm(cbind(infected, total) ~ block*variety + variety*n.trt*status2,
                family = quasibinomial, data = fert2)
summary(glm3way2)

glm2way2 <- glm(cbind(infected, total) ~ block*variety + (variety+n.trt+status2)^2,
                family = quasibinomial, data = fert2)
summary(glm2way2)

@@

<<glmm1,eval=FALSE>>=
glmm3way <- glmer(cbind(infected, total) ~ variety*n.trt*status2 + (1|block/variety),
                  control = glmerControl(optimizer = "bobyqa"),
                  family = binomial, data = fert.sim1)
@

@@

<<glmmhide1,echo=FALSE,cache=TRUE>>=
glmm3way <- glmer(cbind(infected, total) ~ variety*n.trt*status2 + (1|block/variety),
                  control = glmerControl(optimizer = "bobyqa"),
                  family = binomial, data = fert.sim1)
@

@@

<<glmm2,eval=FALSE>>=
glmm3way <- glmer(cbind(infected, total) ~ variety*n.trt*status2 + (1|block/variety),
                  control = glmerControl(optimizer = "bobyqa",
                                         optCtrl = list("maxfun" = 20000)),
                  family = binomial, data = fert.sim1)
@

@@

<<glmmhide2,echo=FALSE,cache=TRUE>>=
glmmPlot <- glmer(cbind(infected, total) ~ variety*n.trt*status2 + (1|plot),
                  control = glmerControl(optimizer = "bobyqa",
                                         optCtrl = list("maxfun" = 20000)),
                  family = binomial, data = fert2)
glmmVar <- glmer(cbind(infected, total) ~ variety*n.trt*status2 +
                   (1|block/variety),
                 control = glmerControl(optimizer = "bobyqa",
                                        optCtrl = list("maxfun" = 20000)),
                 family = binomial, data = fert2)
@

@@

glmBin <- glm(cbind(infected, total) ~ block*variety + variety*n.trt*status2,
              family = binomial, data = fert2)
# Warning message:
# glm.fit: fitted probabilities numerically 0 or 1 occurred
# Residual deviance: 427.939  on 125  degrees of freedom
# 427.939 / 125 = 3.423512

glmQBin <- glm(cbind(infected, total) ~ block*variety + variety*n.trt*status2,
               family = quasibinomial, data = fert2)
# No warnings
# Residual deviance: 427.939  on 125  degrees of freedom
# (Dispersion parameter for quasibinomial family taken to be 3.2817219)

glmmVar <- glmer(cbind(infected, total) ~ variety*n.trt*status2 +
                   (1|block/variety),
                 control = glmerControl(optimizer = "bobyqa",
                                        optCtrl = list("maxfun" = 20000)),
                 family = binomial, data = fert2)
# Warning messages:
# 1: In checkConv(attr(opt, "derivs"), opt$par, ctrl = control$checkConv,  :
#   unable to evaluate scaled gradient
# 2: In checkConv(attr(opt, "derivs"), opt$par, ctrl = control$checkConv,  :
#   Model failed to converge: degenerate  Hessian with 2 negative eigenvalues
# variety:block (Intercept) 0.46633  0.68289
# block         (Intercept) 0.00000  0.00000
# varietySNMS:n.trtLATE SPRING               -18.53042 2535.76041
# varietySNMS:status2CNTL                    -18.50610 2527.55827
# varietySNMS:n.trtEARLY SPRING:status2CNTL   19.11619 2527.55834
# varietySNMS:n.trtLATE SPRING:status2CNTL    19.07429 4381.94303
# deviance df.resid
#    865.7      148
# 865.7 / 148 = 5.8493243

glmmVarPlot <- glmer(cbind(infected, total) ~ variety*n.trt*status2 +
                       (1|block/variety) + (1|plot),
                     control = glmerControl(optimizer = "bobyqa",
                                            optCtrl = list("maxfun" = 20000)),
                     family = binomial, data = fert2)
# Warning messages:
# 1: In checkConv(attr(opt, "derivs"), opt$par, ctrl = control$checkConv,  :
#   unable to evaluate scaled gradient
# 2: In checkConv(attr(opt, "derivs"), opt$par, ctrl = control$checkConv,  :
#   Model failed to converge: degenerate  Hessian with 2 negative eigenvalues# plot          (Intercept) 1.03818  1.0189
# variety:block (Intercept) 0.28324  0.5322
# block         (Intercept) 0.00000  0.0000
# varietySNMS:n.trtLATE SPRING                -20.10372  8008.61084
# varietySNMS:status2CNTL                     -19.96154  7604.01229
# varietySNMS:n.trtEARLY SPRING:status2CNTL    20.93076  7604.01238
# varietySNMS:n.trtLATE SPRING:status2CNTL     20.61822 13745.14025
# deviance df.resid
#    726.8      147
# 726.8 / 147 = 4.9442177

glmmRow <- glmer(cbind(infected, total) ~ variety*n.trt*status2 +
                   (1|block/row),
                 control = glmerControl(optimizer = "bobyqa",
                                        optCtrl = list("maxfun" = 20000)),
                 family = binomial, data = fert2)
# Warning messages:
# 1: In checkConv(attr(opt, "derivs"), opt$par, ctrl = control$checkConv,  :
#   unable to evaluate scaled gradient
# 2: In checkConv(attr(opt, "derivs"), opt$par, ctrl = control$checkConv,  :
#   Model failed to converge: degenerate  Hessian with 3 negative eigenvalues
# row:block (Intercept) 0.46633  0.68289
# block     (Intercept) 0.00000  0.00000
# varietySNMS:n.trtLATE SPRING               -18.52017 2522.79684
# varietySNMS:status2CNTL                    -18.50714 2528.86662
# varietySNMS:n.trtEARLY SPRING:status2CNTL   19.11723 2528.86668
# varietySNMS:n.trtLATE SPRING:status2CNTL    19.06600 4374.54075
# deviance df.resid
#    865.7      148
# 865.7 / 148 = 5.8493243

glmmRowPlot <- glmer(cbind(infected, total) ~ variety*n.trt*status2 +
                       (1|block/row) + (1|plot),
                     control = glmerControl(optimizer = "bobyqa",
                                            optCtrl = list("maxfun" = 20000)),
                     family = binomial, data = fert2)
# Warning messages:
# 1: In checkConv(attr(opt, "derivs"), opt$par, ctrl = control$checkConv,  :
#   unable to evaluate scaled gradient
# 2: In checkConv(attr(opt, "derivs"), opt$par, ctrl = control$checkConv,  :
#   Model failed to converge: degenerate  Hessian with 2 negative eigenvalues
# plot      (Intercept) 1.03818  1.0189
# row:block (Intercept) 0.28324  0.5322
# block     (Intercept) 0.00000  0.0000
# varietySNMS:n.trtLATE SPRING                -20.06813  7867.38080
# varietySNMS:status2CNTL                     -19.95030  7561.42603
# varietySNMS:n.trtEARLY SPRING:status2CNTL    20.91952  7561.42612
# varietySNMS:n.trtLATE SPRING:status2CNTL     20.67575 13394.17908
# deviance df.resid
#    726.8      147
# 726.8 / 147 = 4.9442177

glmmPlot <- glmer(cbind(infected, total) ~ variety*n.trt*status2 + (1|plot),
                  control = glmerControl(optimizer = "bobyqa",
                                         optCtrl = list("maxfun" = 20000)),
                  family = binomial, data = fert2)
# Warning message:
# In checkConv(attr(opt, "derivs"), opt$par, ctrl = control$checkConv,  :
#   Model is nearly unidentifiable: large eigenvalue ratio
#  - Rescale variables?
# plot   (Intercept) 1.3056   1.1426
# varietySNMS:n.trtLATE SPRING              -19.20256  371.85490
# varietySNMS:status2CNTL                   -19.21857  211.30941
# varietySNMS:n.trtEARLY SPRING:status2CNTL  20.44165  211.30904
# varietySNMS:n.trtLATE SPRING:status2CNTL   19.91369  260.18702
# deviance df.resid
#    731.7      149
# 731.7 / 149 = 4.9107383

glmmVarS <- glmer(cbind(infected, total) ~ variety*n.trt*status2 +
                    (variety|block),
                  control = glmerControl(optimizer = "bobyqa",
                                         optCtrl = list("maxfun" = 20000)),
                  family = binomial, data = fert2)
# Warning messages:
# 1: In checkConv(attr(opt, "derivs"), opt$par, ctrl = control$checkConv,  :
#   unable to evaluate scaled gradient
# 2: In checkConv(attr(opt, "derivs"), opt$par, ctrl = control$checkConv,  :
#   Model failed to converge: degenerate  Hessian with 3 negative eigenvalues
# block  (Intercept) 0.62664  0.79160
#        varietyPRHN 0.98385  0.99189  -0.933
#        varietySNMS 3.09082  1.75807  -0.713  0.876
#        varietyTAM  0.38672  0.62186  -0.471  0.432  0.162
#        varietyYSTN 0.86358  0.92929  -0.798  0.687  0.277  0.806
# varietySNMS:n.trtLATE SPRING               -18.54752 2546.52186
# varietySNMS:status2CNTL                    -18.51916 2537.53981
# varietySNMS:n.trtEARLY SPRING:status2CNTL   19.13402 2537.53988
# varietySNMS:n.trtLATE SPRING:status2CNTL    19.08471 4406.80309
# deviance df.resid
#    842.8      135
# 842.8 / 135 = 6.2422222

@@

noSNMS3way <- glmer(cbind(infected, total) ~ variety*n.trt*status2 +
                    (1|block/row), subset = variety != "SNMS",
                    control = glmerControl(optimizer = "bobyqa"),
                    family = binomial, data = fert2)
summary(noSNMS3way)
Anova(noSNMS3way)

noSNMS2way <- glmer(cbind(infected, total) ~ variety*n.trt + variety*status2 +
                    n.trt*status2 + (1|block/row), subset = variety != "SNMS",
                    control = glmerControl(optimizer = "bobyqa"),
                    family = binomial, data = fert2)
summary(noSNMS2way)
Anova(noSNMS2way)

@@

\noindent The Anova table for the three-way interaction model is shown below (Table \ref{3way}). This table should only be used to assess evidence for the three-way interaction term. We recommend the following wording,
\begin{quote}
There is no evidence that the interaction between variety and nitrogen application time differs between inoculated and control plots (p-value$=0.388$ from F-stat$=1.07$ on $8$ and $125$ df).
\end{quote}

@@

<<anovadisplay2, echo=TRUE, results='hide'>>=
glm2way <- glm(cbind(infected, total) ~ block*variety + variety*n.trt
               + variety*status2+n.trt*status2,
               family = quasibinomial, data = fert.sim1)
Anova(glm2way, test="F")
@

@@

\noindent Here is how we would interpret the results in the applicable rows in the above the two-way interaction Anova table (Table \ref{twoway}).
\begin{enumerate}
\item \verb+n.trt:status2+ ``There is no evidence that the relationship between nitrogen application time and the odds of virus infection differs between inoculated and control plots after controlling for all other two way interactions and main effects in the model (p-value$=0.380$ from F-stat$=0.973$ on $2$ and $133$ df).
\item \verb+variety:status2+ ``There is no evidence that the relationship between variety and the odds of virus infection differs between inoculated and control plots after controlling for all other two way interactions and main effects in the model (p-value$=0.304$ from F-stat$=1.22$ on $4$ and $133$ df).
\item \verb+variety:n.trt+ ``There is no evidence that the relationship between nitrogen application time and the odds of virus infection depends on variety after controlling for all other two way interactions and main effects in the model (p-value$=0.478$ from F-stat$=0.950$ on $8$ and $133$ df).
\end{enumerate}

\noindent If you remove another term from the model, refit the simpler model and reassess evidence for each effect. For example, suppose we remove the variety by inoculation status interaction term (Table \ref{remove}). Our assessment of the variety by nitrogen application interaction effect in the following simpler model would be,
\begin{quote}
There is no evidence that the relationship between nitrogen application time and the odds of virus infection depends on variety after controlling for all main effects and the two way interaction between variety and inoculation status (p-value$=0.4536$ from F-stat$=0.98$ on $8$ and $135$ df).
\end{quote}
<<refitno2way, echo=FALSE, results='asis'>>=
glm2way2 <- glm(cbind(infected, total) ~ block*variety + variety*n.trt
               + variety*status2,
               family = quasibinomial, data = fert.sim1)
print(xtable(Anova(glm2way2, test="F"), caption="Anova table for the model with two way interactions between variety and nitrogen and variety and inoculation status.", label="remove"), caption.placement = "top", size='footnotesize', table.placement = getOption("xtable.table.placement", "H"))
@

<<refitno2way3, echo=FALSE, results='asis', include=FALSE>>=
glm2way3 <- glm(cbind(infected, total) ~ block*variety + variety*n.trt,
               family = quasibinomial, data = fert.sim1)
print(xtable(Anova(glm2way3, test="F"), caption="Anova table for the model with a two way interaction between variety and nitrogen application time."), caption.placement = "top", table.placement = getOption("xtable.table.placement", "H"))
@

@@

\noindent To assess evidence for the \verb+n.trt+ effect in the simplest recommended model, we would say,
\begin{quote}
There is no evidence that the odds of virus infection depends on nitrogen application time after controlling for block, variety, inoculation status, and the $block*variety$ interaction (p-value$=0.612$ from F-stat$=0.49$ on $2$ and $147$ df).
\end{quote}

<<glmnointeractions, echo=TRUE, results='hide'>>=
glmnointeractions <- glm(cbind(infected, total) ~ block*variety + variety + n.trt + status2,
               family = quasibinomial, data = fert.sim1)
Anova(glmnointeractions, test="F")
@

@@

<<dropindevftest, echo=FALSE>>=
glm.no3way <- glm(cbind(infected, total) ~ block*variety + variety*n.trt + n.trt*status2 + variety*status2,
               family = quasibinomial, data = fert.sim1)
#drop in dev F-test
#((196-185)/8)/(185/125)
#0.929
#1-pf(0.929, 11, 125)
#0.512
@

@@

\noindent For these data, we would describe the nitrogen application effect for inoculated plants of variety $1$ as follows:
\begin{quote}
For inoculated plants of variety $1$, the odds of virus infection when nitrogen is applied in the early spring is estimated to be $3.01$ times the odds of virus infection when nitrogen is applied in the fall, with a $95\%$ likelihood-based confidence interval from  $0.295$ to $104.9$ times. For inoculated plants of variety $1$, the odds of virus infection when nitrogen is applied in the late spring is estimated to be $5.05$ times the odds of virus infection when nitrogen is applied in the fall, with a $95\%$ likelihood-based confidence interval from $0.652$ to $166.4$ times.
\end{quote}

<<second, echo=TRUE, message=FALSE, eval=FALSE>>=
#suppose we want to the effects of introgen application for
#inoculated plants of variety 3
#first relevel
var3 <- relevel(fert.sim1$variety, "Var3")
#then fit model with var3
glm.refvar3 <- glm(cbind(infected, total) ~ block/var3 + var3*n.trt*status2,
               family = quasibinomial, data = fert.sim1)
#interpret results
exp(coef(glm.refvar3)[11])
exp(confint(glm.refvar3, 11))
exp(coef(glm.refvar3)[12])
exp(confint(glm.refvar3, 12))

#now suppose we want to the effects of introgen application for
#non-inoculated plants of variety 3
#first relevel the status variable
statuscntl <- relevel(fert.sim1$status2, "CNTL")
#then fit model with var3
glm.refvar3cntl <- glm(cbind(infected, total) ~ block/var3 + var3*n.trt*statuscntl,
               family = quasibinomial, data = fert.sim1)
#interpret results
exp(coef(glm.refvar3cntl)[11])
exp(confint(glm.refvar3cntl, 11))
exp(coef(glm.refvar3cntl)[12])
exp(confint(glm.refvar3cntl, 12))
@

>>>>

<<fakedata,echo=FALSE>>=
# Create some fake data with no differences anywhere
set.seed(983724)
fert.sim1 <- data.frame(expand.grid(
  "status" = c("CNTL", "INOC"),
  "n.application" = c("FALL", "EARLY SPRING", "LATE SPRING"),
  "variety" = c("Var1", "Var2", "Var3", "Var4", "Var5"),
  "block" = LETTERS[1:6],
  "total" = 30),
  "infected" = rnbinom(180, 4, 7/9))
#  "infected" = sample(fert2$infected, 180, replace = TRUE))
fert.sim1$uninfected <- with(fert.sim1, total - infected)
@

@@

<<comparison, echo=FALSE, message=FALSE, results='asis'>>=
glmQBin.noSNMS <- glm(cbind(infected, uninfected) ~ block*variety +
                        variety*n.trt*status2,
               family = quasibinomial, data = subset(fert2, variety != "SNMS"))
print(xtable(summary(glmQBin.noSNMS)$coefficients[39:44,], caption="Estimates, standard errors, and tests for three way interaction coefficients in the quasibinomial model.", label="qbin"), caption.placement="top", size="footnotesize", table.placement = getOption("xtable.table.placement", "H"))

noSNMS3way <- glmer(cbind(infected, uninfected) ~ variety*n.trt*status2 +
                    (1|block/row), subset = variety != "SNMS",
                    control = glmerControl(optimizer = "bobyqa"),
                    family = binomial, data = fert2)

@@

<<startingmodel,warning=FALSE,results='hide',cache=TRUE>>=
glm3way <- glm(cbind(infected, uninfected) ~ block*variety + variety*n.trt*status2,
               family = quasibinomial, data = fert.sim1)
@

@@

<<novar1,warning=FALSE,results='hide',cache=TRUE>>=
noVar1 <- glm(cbind(infected, uninfected) ~ block*variety + variety*n.trt*status2,
              family = quasibinomial, data = fert.sim1, subset = variety!="Var1")
@

@@

<<separatemodels,warning=FALSE,results='hide',cache=TRUE>>=
onlyINOC <- glm(cbind(infected, uninfected) ~ block*variety + variety*n.trt,
                family = quasibinomial, data = fert.sim1, subset = status=="INOC")

onlyCNTL <- glm(cbind(infected, uninfected) ~ block*variety + variety*n.trt,
                family = quasibinomial, data = fert.sim1, subset = status=="CNTL")
@

<<models5,echo=FALSE,warning=FALSE,message=FALSE,results='hide',cache=TRUE>>=
glm3way2 <- glm(cbind(infected, uninfected) ~ block*variety + variety*n.trt*status2,
                family = quasibinomial, data = fert2)
summary(glm3way2)

glm2way2 <- glm(cbind(infected, uninfected) ~ block*variety + (variety+n.trt+status2)^2,
                family = quasibinomial, data = fert2)
summary(glm2way2)

@@

<<glmm1,eval=FALSE>>=
glmm3way <- glmer(cbind(infected, uninfected) ~ variety*n.trt*status2 + (1|block/variety),
                  control = glmerControl(optimizer = "bobyqa"),
                  family = binomial, data = fert.sim1)
@

@@

<<glmmhide1,echo=FALSE,cache=TRUE>>=
glmm3way <- glmer(cbind(infected, uninfected) ~ variety*n.trt*status2 + (1|block/variety),
                  control = glmerControl(optimizer = "bobyqa"),
                  family = binomial, data = fert.sim1)
@

@@

<<glmm2,eval=FALSE>>=
glmm3way <- glmer(cbind(infected, uninfected) ~ variety*n.trt*status2 + (1|block/variety),
                  control = glmerControl(optimizer = "bobyqa",
                                         optCtrl = list("maxfun" = 20000)),
                  family = binomial, data = fert.sim1)
@

@@

<<glmmhide2,echo=FALSE,cache=TRUE>>=
glmmPlot <- glmer(cbind(infected, uninfected) ~ variety*n.trt*status2 + (1|plot),
                  control = glmerControl(optimizer = "bobyqa",
                                         optCtrl = list("maxfun" = 20000)),
                  family = binomial, data = fert2)
glmmVar <- glmer(cbind(infected, uninfected) ~ variety*n.trt*status2 +
                   (1|block/variety),
                 control = glmerControl(optimizer = "bobyqa",
                                        optCtrl = list("maxfun" = 20000)),
                 family = binomial, data = fert2)
@

@@

glmBin <- glm(cbind(infected, uninfected) ~ block*variety + variety*n.trt*status2,
              family = binomial, data = fert2)
# Warning message:
# glm.fit: fitted probabilities numerically 0 or 1 occurred
# Residual deviance: 427.939  on 125  degrees of freedom
# 427.939 / 125 = 3.423512

glmQBin <- glm(cbind(infected, uninfected) ~ block*variety + variety*n.trt*status2,
               family = quasibinomial, data = fert2)
# No warnings
# Residual deviance: 427.939  on 125  degrees of freedom
# (Dispersion parameter for quasibinomial family taken to be 3.2817219)

glmmVar <- glmer(cbind(infected, uninfected) ~ variety*n.trt*status2 +
                   (1|block/variety),
                 control = glmerControl(optimizer = "bobyqa",
                                        optCtrl = list("maxfun" = 20000)),
                 family = binomial, data = fert2)
# Warning messages:
# 1: In checkConv(attr(opt, "derivs"), opt$par, ctrl = control$checkConv,  :
#   unable to evaluate scaled gradient
# 2: In checkConv(attr(opt, "derivs"), opt$par, ctrl = control$checkConv,  :
#   Model failed to converge: degenerate  Hessian with 2 negative eigenvalues
# variety:block (Intercept) 0.46633  0.68289
# block         (Intercept) 0.00000  0.00000
# varietySNMS:n.trtLATE SPRING               -18.53042 2535.76041
# varietySNMS:status2CNTL                    -18.50610 2527.55827
# varietySNMS:n.trtEARLY SPRING:status2CNTL   19.11619 2527.55834
# varietySNMS:n.trtLATE SPRING:status2CNTL    19.07429 4381.94303
# deviance df.resid
#    865.7      148
# 865.7 / 148 = 5.8493243

glmmVarPlot <- glmer(cbind(infected, uninfected) ~ variety*n.trt*status2 +
                       (1|block/variety) + (1|plot),
                     control = glmerControl(optimizer = "bobyqa",
                                            optCtrl = list("maxfun" = 20000)),
                     family = binomial, data = fert2)
# Warning messages:
# 1: In checkConv(attr(opt, "derivs"), opt$par, ctrl = control$checkConv,  :
#   unable to evaluate scaled gradient
# 2: In checkConv(attr(opt, "derivs"), opt$par, ctrl = control$checkConv,  :
#   Model failed to converge: degenerate  Hessian with 2 negative eigenvalues# plot          (Intercept) 1.03818  1.0189
# variety:block (Intercept) 0.28324  0.5322
# block         (Intercept) 0.00000  0.0000
# varietySNMS:n.trtLATE SPRING                -20.10372  8008.61084
# varietySNMS:status2CNTL                     -19.96154  7604.01229
# varietySNMS:n.trtEARLY SPRING:status2CNTL    20.93076  7604.01238
# varietySNMS:n.trtLATE SPRING:status2CNTL     20.61822 13745.14025
# deviance df.resid
#    726.8      147
# 726.8 / 147 = 4.9442177

glmmRow <- glmer(cbind(infected, uninfected) ~ variety*n.trt*status2 +
                   (1|block/row),
                 control = glmerControl(optimizer = "bobyqa",
                                        optCtrl = list("maxfun" = 20000)),
                 family = binomial, data = fert2)
# Warning messages:
# 1: In checkConv(attr(opt, "derivs"), opt$par, ctrl = control$checkConv,  :
#   unable to evaluate scaled gradient
# 2: In checkConv(attr(opt, "derivs"), opt$par, ctrl = control$checkConv,  :
#   Model failed to converge: degenerate  Hessian with 3 negative eigenvalues
# row:block (Intercept) 0.46633  0.68289
# block     (Intercept) 0.00000  0.00000
# varietySNMS:n.trtLATE SPRING               -18.52017 2522.79684
# varietySNMS:status2CNTL                    -18.50714 2528.86662
# varietySNMS:n.trtEARLY SPRING:status2CNTL   19.11723 2528.86668
# varietySNMS:n.trtLATE SPRING:status2CNTL    19.06600 4374.54075
# deviance df.resid
#    865.7      148
# 865.7 / 148 = 5.8493243

glmmRowPlot <- glmer(cbind(infected, uninfected) ~ variety*n.trt*status2 +
                       (1|block/row) + (1|plot),
                     control = glmerControl(optimizer = "bobyqa",
                                            optCtrl = list("maxfun" = 20000)),
                     family = binomial, data = fert2)
# Warning messages:
# 1: In checkConv(attr(opt, "derivs"), opt$par, ctrl = control$checkConv,  :
#   unable to evaluate scaled gradient
# 2: In checkConv(attr(opt, "derivs"), opt$par, ctrl = control$checkConv,  :
#   Model failed to converge: degenerate  Hessian with 2 negative eigenvalues
# plot      (Intercept) 1.03818  1.0189
# row:block (Intercept) 0.28324  0.5322
# block     (Intercept) 0.00000  0.0000
# varietySNMS:n.trtLATE SPRING                -20.06813  7867.38080
# varietySNMS:status2CNTL                     -19.95030  7561.42603
# varietySNMS:n.trtEARLY SPRING:status2CNTL    20.91952  7561.42612
# varietySNMS:n.trtLATE SPRING:status2CNTL     20.67575 13394.17908
# deviance df.resid
#    726.8      147
# 726.8 / 147 = 4.9442177

glmmPlot <- glmer(cbind(infected, uninfected) ~ variety*n.trt*status2 + (1|plot),
                  control = glmerControl(optimizer = "bobyqa",
                                         optCtrl = list("maxfun" = 20000)),
                  family = binomial, data = fert2)
# Warning message:
# In checkConv(attr(opt, "derivs"), opt$par, ctrl = control$checkConv,  :
#   Model is nearly unidentifiable: large eigenvalue ratio
#  - Rescale variables?
# plot   (Intercept) 1.3056   1.1426
# varietySNMS:n.trtLATE SPRING              -19.20256  371.85490
# varietySNMS:status2CNTL                   -19.21857  211.30941
# varietySNMS:n.trtEARLY SPRING:status2CNTL  20.44165  211.30904
# varietySNMS:n.trtLATE SPRING:status2CNTL   19.91369  260.18702
# deviance df.resid
#    731.7      149
# 731.7 / 149 = 4.9107383

glmmVarS <- glmer(cbind(infected, uninfected) ~ variety*n.trt*status2 +
                    (variety|block),
                  control = glmerControl(optimizer = "bobyqa",
                                         optCtrl = list("maxfun" = 20000)),
                  family = binomial, data = fert2)
# Warning messages:
# 1: In checkConv(attr(opt, "derivs"), opt$par, ctrl = control$checkConv,  :
#   unable to evaluate scaled gradient
# 2: In checkConv(attr(opt, "derivs"), opt$par, ctrl = control$checkConv,  :
#   Model failed to converge: degenerate  Hessian with 3 negative eigenvalues
# block  (Intercept) 0.62664  0.79160
#        varietyPRHN 0.98385  0.99189  -0.933
#        varietySNMS 3.09082  1.75807  -0.713  0.876
#        varietyTAM  0.38672  0.62186  -0.471  0.432  0.162
#        varietyYSTN 0.86358  0.92929  -0.798  0.687  0.277  0.806
# varietySNMS:n.trtLATE SPRING               -18.54752 2546.52186
# varietySNMS:status2CNTL                    -18.51916 2537.53981
# varietySNMS:n.trtEARLY SPRING:status2CNTL   19.13402 2537.53988
# varietySNMS:n.trtLATE SPRING:status2CNTL    19.08471 4406.80309
# deviance df.resid
#    842.8      135
# 842.8 / 135 = 6.2422222

@@

noSNMS3way <- glmer(cbind(infected, uninfected) ~ variety*n.trt*status2 +
                    (1|block/row), subset = variety != "SNMS",
                    control = glmerControl(optimizer = "bobyqa"),
                    family = binomial, data = fert2)
summary(noSNMS3way)
Anova(noSNMS3way)

noSNMS2way <- glmer(cbind(infected, uninfected) ~ variety*n.trt + variety*status2 +
                    n.trt*status2 + (1|block/row), subset = variety != "SNMS",
                    control = glmerControl(optimizer = "bobyqa"),
                    family = binomial, data = fert2)
summary(noSNMS2way)
Anova(noSNMS2way)

@@

\noindent The Anova table for the three-way interaction model is shown below (Table \ref{3way}). This table should only be used to assess evidence for the three-way interaction term. We recommend the following wording,
\begin{quote}
There is no evidence that the interaction between variety and nitrogen application time differs between inoculated and control plots (p-value$=0.406$ from F-stat$=1.05$ on $8$ and $125$ df).
\end{quote}

@@

<<anovadisplay2, echo=TRUE, results='hide'>>=
glm2way <- glm(cbind(infected, uninfected) ~ block*variety + variety*n.trt
               + variety*status2+n.trt*status2,
               family = quasibinomial, data = fert.sim1)
Anova(glm2way, test="F")
@

@@

\noindent Here is how we would interpret the results in the applicable rows in the above the two-way interaction Anova table (Table \ref{twoway}).
\begin{enumerate}
\item \verb+n.trt:status2+ ``There is no evidence that the relationship between nitrogen application time and the odds of virus infection differs between inoculated and control plots after controlling for all other two way interactions and main effects in the model (p-value$=0.386$ from F-stat$=0.959$ on $2$ and $133$ df).
\item \verb+variety:status2+ ``There is no evidence that the relationship between variety and the odds of virus infection differs between inoculated and control plots after controlling for all other two way interactions and main effects in the model (p-value$=0.302$ from F-stat$=1.23$ on $4$ and $133$ df).
\item \verb+variety:n.trt+ ``There is no evidence that the relationship between nitrogen application time and the odds of virus infection depends on variety after controlling for all other two way interactions and main effects in the model (p-value$=0.465$ from F-stat$=0.967$ on $8$ and $133$ df).
\end{enumerate}

\noindent If you remove another term from the model, refit the simpler model and reassess evidence for each effect. For example, suppose we remove the variety by inoculation status interaction term (Table \ref{remove}). Our assessment of the variety by nitrogen application interaction effect in the following simpler model would be,
\begin{quote}
There is no evidence that the relationship between nitrogen application time and the odds of virus infection depends on variety after controlling for all main effects and the two way interaction between variety and inoculation status (p-value$=0.440$ from F-stat$=0.999$ on $8$ and $135$ df).
\end{quote}
<<refitno2way, echo=FALSE, results='asis'>>=
glm2way2 <- glm(cbind(infected, uninfected) ~ block*variety + variety*n.trt
               + variety*status2,
               family = quasibinomial, data = fert.sim1)
print(xtable(Anova(glm2way2, test="F"), caption="Anova table for the model with two way interactions between variety and nitrogen and variety and inoculation status.", label="remove"), caption.placement = "top", size='footnotesize', table.placement = getOption("xtable.table.placement", "H"))
@

<<refitno2way3, echo=FALSE, results='asis', include=FALSE>>=
glm2way3 <- glm(cbind(infected, uninfected) ~ block*variety + variety*n.trt,
               family = quasibinomial, data = fert.sim1)
print(xtable(Anova(glm2way3, test="F"), caption="Anova table for the model with a two way interaction between variety and nitrogen application time."), caption.placement = "top", table.placement = getOption("xtable.table.placement", "H"))
@

@@

\noindent To assess evidence for the \verb+n.trt+ effect in the simplest recommended model, we would say,
\begin{quote}
There is no evidence that the odds of virus infection depends on nitrogen application time after controlling for block, variety, inoculation status, and the $block*variety$ interaction (p-value$=0.633$ from F-stat$=0.459$ on $2$ and $147$ df).
\end{quote}

<<glmnointeractions, echo=TRUE, results='hide'>>=
glmnointeractions <- glm(cbind(infected, uninfected) ~ block*variety + variety + n.trt + status2,
               family = quasibinomial, data = fert.sim1)
Anova(glmnointeractions, test="F")
@

@@

<<dropindevftest, echo=FALSE>>=
glm.no3way <- glm(cbind(infected, uninfected) ~ block*variety + variety*n.trt + n.trt*status2 + variety*status2,
               family = quasibinomial, data = fert.sim1)
#drop in dev F-test
#((196-185)/8)/(185/125)
#0.929
#1-pf(0.929, 11, 125)
#0.512
@

@@

\noindent For these data, we would describe the nitrogen application effect for inoculated plants of variety $1$ as follows:
\begin{quote}
For inoculated plants of variety $1$, the odds of virus infection when nitrogen is applied in the early spring is estimated to be $3.04$ times the odds of virus infection when nitrogen is applied in the fall, with a $95\%$ likelihood-based confidence interval from  $0.264$ to $136.6$ times. For inoculated plants of variety $1$, the odds of virus infection when nitrogen is applied in the late spring is estimated to be $5.13$ times the odds of virus infection when nitrogen is applied in the fall, with a $95\%$ likelihood-based confidence interval from $0.604$ to $218.4$ times.
\end{quote}

<<second, echo=TRUE, message=FALSE, eval=FALSE>>=
#suppose we want to the effects of introgen application for
#inoculated plants of variety 3
#first relevel
var3 <- relevel(fert.sim1$variety, "Var3")
#then fit model with var3
glm.refvar3 <- glm(cbind(infected, uninfected) ~ block/var3 + var3*n.trt*status2,
               family = quasibinomial, data = fert.sim1)
#interpret results
exp(coef(glm.refvar3)[11])
exp(confint(glm.refvar3, 11))
exp(coef(glm.refvar3)[12])
exp(confint(glm.refvar3, 12))

#now suppose we want to the effects of introgen application for
#non-inoculated plants of variety 3
#first relevel the status variable
statuscntl <- relevel(fert.sim1$status2, "CNTL")
#then fit model with var3
glm.refvar3cntl <- glm(cbind(infected, uninfected) ~ block/var3 + var3*n.trt*statuscntl,
               family = quasibinomial, data = fert.sim1)
#interpret results
exp(coef(glm.refvar3cntl)[11])
exp(confint(glm.refvar3cntl, 11))
exp(coef(glm.refvar3cntl)[12])
exp(confint(glm.refvar3cntl, 12))
@
